TBD

### 2023-03-05 outputs/paper/paper3.qmd
Added To Do section to keep track of tasks
data exploration for financial satisfaction & social status

#### Workspace set-up ####

@@

rm(cleaned_respondent_info)
rm(cleaned_finance_survey)

@@

TBD

@@

## Explore 2021 data

survey_data_2016 =
  combined_data |>
  filter(year == 2016)

#survey_data_2021 = merge(
#  filter(cleaned_respondent_info, year == 2021),
#  filter(cleaned_finance_survey, year == 2021),
#  by = c("year", "id")
#)
```

```{r}
#| message: false
#| echo: false
#| 
survey_data_2021$financial_satisfaction =
  factor(survey_data_2021$financial_satisfaction,
         levels = c(
           "Not Satisfied At All",
           "More or Less Satisfied",
           "Pretty Well Satisfied"
          )
        )

survey_data_2021$gender =
  factor(survey_data_2021$gender,
         levels = c(
           "Male",
           "Female"
          )
        )

survey_data_2021$degree =
  factor(survey_data_2021$degree,
         levels = c(
           "Less Than High School",
           "High School",
           "Associate/Junior College",
           "Bachelors",
           "Graduate"
          )
        )

survey_data_2021$age_band =
  factor(survey_data_2021$age_band,
         levels = c(
           "18-29",
           "30-39",
           "40-49",
           "50-64",
           "65-79",
           ">=80"
          )
        )

survey_data_2021$social_class =
  factor(survey_data_2021$social_class,
         levels = c(
           "Lower Class",
           "Working Class",
           "Lower Middle Class",
           "Middle Class",
           "Upper Middle Class",
           "Upper Class"
          )
        )

survey_data_2021$financial_compare =
  factor(survey_data_2021$financial_compare,
         levels = c(
           "Far Below Average",
           "Below Average",
           "Average",
           "Above Average",
           "Far Above Average"
         ))

survey_data_2021$financial_change =
  factor(survey_data_2021$financial_change,
         levels = c(
           "Getting Worse",
           "Stayed the Same",
           "Getting Better"
         ))

survey_data_2016$financial_change =
  factor(survey_data_2016$financial_change,
         levels = c(
           "Getting Worse",
           "Stayed the Same",
           "Getting Better"
         ))
```

@@

Overall spread

@@

@@

@@

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2016 |>
  filter(!is.na(social_class), !is.na(financial_change)) |>
  ggplot(aes(x = financial_change, group = social_class)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Change") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~social_class) 
```

financial_satisfaction_across_years =
  combined_data |>
    select (year, financial_satisfaction) |>
    group_by(year) |>
    count(financial_satisfaction)

financial_satisfaction_across_years
```

>>>>

#### Workspace set-up and read in data####

@@

rm(cleaned_respondent_info)
rm(cleaned_finance_survey)
```

#### Add factor defintions to define orders for columns in graphs ####

combined_data$financial_satisfaction =
  factor(combined_data$financial_satisfaction,
         levels = c(
           "Not Satisfied At All",
           "More or Less Satisfied",
           "Pretty Well Satisfied"
          )
        )

combined_data$gender =
  factor(combined_data$gender,
         levels = c(
           "Male",
           "Female"
          )
        )

combined_data$degree =
  factor(combined_data$degree,
         levels = c(
           "Less Than High School",
           "High School",
           "Associate/Junior College",
           "Bachelors",
           "Graduate"
          )
        )

combined_data$age_band =
  factor(combined_data$age_band,
         levels = c(
           "18-29",
           "30-39",
           "40-49",
           "50-64",
           "65-79",
           ">=80"
          )
        )

combined_data$social_class =
  factor(combined_data$social_class,
         levels = c(
           "Lower Class",
           "Working Class",
           "Middle Class",
           "Upper Class"
          )
        )

combined_data$social_class1 =
  factor(combined_data$social_class1,
         levels = c(
           "Lower Class",
           "Working Class",
           "Lower Middle Class",
           "Middle Class",
           "Upper Middle Class",
           "Upper Class"
          )
        )

combined_data$cohort_band =
  factor(combined_data$cohort_band,
         levels = c(
           "Greatest",
           "Silent",
           "Boomer",
           "Gen X",
           "Millenials",
           "Gen Z"
           )
         )

combined_data$financial_compare =
  factor(combined_data$financial_compare,
         levels = c(
           "Far Below Average",
           "Below Average",
           "Average",
           "Above Average",
           "Far Above Average"
         ))

combined_data$financial_change =
  factor(combined_data$financial_change,
         levels = c(
           "Getting Worse",
           "Stayed the Same",
           "Getting Better"
         ))
```

## Save a slice of 2021 and 2016 data

survey_data_2016 =
  combined_data |>
  filter(year == 2016)
```

## To Do

Christina

-   financial satisfaction

-   Social status

-   calculating % ... over the year for graph

-   Looking into social class getting it over the years - DONE

Michaela

-   financial change

-   financial compare

Other

-   Supplementary survey

-   Simulated data

-   Data validation tests

@@

The rich getting richer; poor getting poorer

The super upper class are in a bubble and delusional

Female and male differences

@@

@@

Overall spread in 2021

@@

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~social_class1)
```

@@

Financial Satisfaction Over the Years

## Create variable to track financial satisfaction across years ##

financial_satisfaction_across_years =
  combined_data |>
    select(year, financial_satisfaction) |>
    group_by(year) |>
    count(financial_satisfaction)
```

Financial satisfaction - 2002, 2004, 2006 years have a lot of NAs!

financial_satisfaction_across_years |>
  filter(!is.na(financial_satisfaction)) |>
  #filter(!(year %in% c(2002, 2004, 2006))) |>
  mutate(percent = n / sum(n)) |>
  ggplot(aes(x = year, y = percent, color = financial_satisfaction)) +
  geom_smooth(method=lm, se=FALSE) + 
  #geom_bar(stat="identity") +
  geom_line() +
  scale_y_continuous(labels=scales::percent)

Financial satisfaction - tracking cohort over time

financial_satisfaction_across_years_by_cohort |>
  filter(!is.na(financial_satisfaction)) |>
  filter(!(is.na(cohort_band))) |>
  mutate(percent = n / sum(n)) |>
  #filter(financial_satisfaction == "Not Satisfied At All") |>
  #mutate(satisfied = 1 - percent) |>
  ggplot(aes(x = year, y = percent, color = financial_satisfaction )) +
  geom_line() +
  geom_smooth(method=lm, se=FALSE) + 
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(vars(cohort_band), nrow = 3, ncol = 3) +
  theme(legend.position="bottom")
```

Social class over the years

## Create variable to track financial satisfaction across years ##

social_class_across_years =
  combined_data |>
  select(year, social_class) |>
  group_by(year) |>
  count(social_class)

social_class_across_years |>
  filter(!is.na(social_class)) |>
  mutate(percent = n / sum(n)) |>
  ggplot(aes(x = year, y = percent, color = social_class)) +
  #geom_bar(stat="identity") +
  geom_line() +
  scale_y_continuous(labels=scales::percent) +
  geom_smooth(method=lm, se=FALSE)

### Financial Compare

@@

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
# survey_data_2016 |>
#   filter(!is.na(social_class), !is.na(financial_change)) |>
#   ggplot(aes(x = financial_change, group = social_class)) +
#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
#   scale_y_continuous(labels=scales::percent) +
#   ylab("Percentage of response") +
#   xlab("Financial Change") +
#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
#   theme(legend.position="none") +
#   facet_grid(~social_class) 
```
